Symbolic Weighted Language Models and\\ Quantitative Parsing over Infinite Alphabets

### 2021-10-12 algo-best-alt.tex
proof algo cont.

%Intuitively, 
\begin{lemma}[Correctness]
For all $q_0, q_1, q_2 \in Q$, $p\in P$, 
and all path $\pi$ of $\G_A$ with $\src(\pi) = \<q_0, q_0>$,
($i$) if $\trg(\pi) = \< q_1, q_2>$, 
then there exists $u \in \Delta^*$ such that 
$\weight_A(q_1[\bot], u, q_1[\bot]) = \weight(\pi)$,
%
($ii$) if $\trg(\pi) = \< q_1, p, q_2>$,
then there exists $v \in \Delta^*$ such that 
${\displaystyle\bigoplus_{c \in \Deltac}}
 \weight_A\bigl(\configup{q_1}{\< c, p> \stackup \top }, v, \configup{q_2}{\<c, p> \stackup \top}\bigr)
 = \weight(\pi)$.
\end{lemma}
%
\begin{proof}
We prove ($i$) and ($ii$) simultaneously by induction on the length of the path $\pi$.
...\qed
\end{proof}

\begin{lemma}[Completeness]
For all $q, q' \in Q$, $p\in P$, 
($i$) there exists a path~$\pi$ of~$\G_A$ 
such that $\src(\pi) = \<q_0, q_0>$ for some $q_0 \in Q$, 
$\trg(\pi) = \< q, q'>$,
and $\weight(\pi) = b_\bot(q, q')$, 
($ii$) there exists a path $\pi'$ of $\G_A$ 
such that $\src(\pi) = \<q_0, q_0>$ for some $q_0 \in Q$, 
$\trg(\pi) = \< q_1, p, q_2>$,
and $\weight(\pi) = \bigoplus_\Deltac b_\top(q_1, p, q_2)$.
\end{lemma}
%
\begin{proof}
By associativity, commutativity and distributivity for $\Semiring$,
\eqref{eq:bbot} can be rewritten into the form, unfolding~\eqref{eq:SWVPA-weight}:
\[
b_\bot(q, q') = 
\bigoplus_{t\in \Delta^*} 
\bigoplus_{\begin{array}{c}
           \scriptstyle q_0,\ldots, q_n \in Q\\
           \scriptstyle p_{i},\ldots, p_{k} \in P
           \end{array}}
\bigotimes_{i=i}^n w_i(\tau_i)           
\]
where $n$ is the length of $t$, $k \leq n$,
$q_0 = q$, $q_n = q'$, 
for all $1 \leq i < n$, 
$w_i$ is one of the functions of $\bar\wei$,
$\tau_i$ is a transition of $A$ and
$\src(\tau_i) = q_{i-1}$, $\snd(\tau_i) = q_i$.
%
Since $\Semiring$ is total, there exists finite sequences as above such that
\( b_\bot(q, q') = \bigotimes_{i=i}^n w_i(\tau_i)\).
%
There might exists, for $q$ and $q'$,
several finite sequences~$\bar\tau$ and~$\bar{w}$ as above, 
let us choose arbitrarily one of minimal length~$n$. 
This integer $n$ will be denoted $n_{q, q'}$ in the following.
%We represent a computation of $A$ on $t$ by a sequence $\rho$ of its transitions,

Similarly, following~\eqref{eq:btop} and~\eqref{eq:SWVPA-weight}, 
$\bigoplus_\Deltac b_\top(q_1, q_2)$ can be put in the form:
\[
{\textstyle\bigoplus_\Deltac} b_\top(q_1, q_2) = 
{\textstyle\bigoplus_\Deltac} \bigoplus_{t\in \Delta^*} 
\bigoplus_{\begin{array}{c}
           \scriptstyle q_0,\ldots, q_n \in Q\\
           \scriptstyle p_{i},\ldots, p_{k} \in P
           \end{array}}
\bigotimes_{i=i}^n w_i(\tau_i)           
\]
and hence 
\( {\textstyle\bigoplus_\Deltac} b_\top(q_1, q_2) = \bigotimes_{i=i}^n w_i(\tau_i) \),
using $c \in \Deltac$ that minimize...
We denote the smallest $n$ as above by $n_{q, p, q'}$.

We show now the existence of a path $\pi$ as expected, by induction on $n$.
%that we can construct a path $\pi$ of weight $b_\bot(q_1, q_2)$
%from $\tau_0,\ldots, \tau_n$ and $w_0,\ldots, w_n$. ...\qed
\end{proof}

>>>>

%Intuitively, 
\begin{lemma}[Correctness]
For all $q_0, q, q' \in Q$, $p\in P$, 
and all path $\pi$ of $\G_A$ with $\src(\pi) = \<q_0, q_0>$,
($i$) if $\trg(\pi) = \< q, q'>$, 
then there exists $u \in \Delta^*$ such that 
$\weight_A(q[\bot], u, q'[\bot]) = \weight(\pi)$,
%
($ii$) if $\trg(\pi) = \< q, p, q'>$,
then there exists $v \in \Delta^*$ such that 
${\displaystyle\bigoplus_{c \in \Deltac}}
 \weight_A\bigl(\configup{q}{\< c, p> \stackup \top }, v, \configup{q'}{\<c, p> \stackup \top}\bigr)
 = \weight(\pi)$.
\end{lemma}
%
\begin{proof}
We prove ($i$) and ($ii$) simultaneously by induction on the length of the path $\pi$.
...\qed
\end{proof}

\begin{lemma}[Completeness] \label{lem:algo-completeness}
For all $q, q' \in Q$, $p\in P$, 
($i$) there exists a path~$\pi$ of~$\G_A$ 
such that $\src(\pi) = \<q_0, q_0>$ for some $q_0 \in Q$, 
$\trg(\pi) = \< q, q'>$,
and $\weight(\pi) = b_\bot(q, q')$, 
($ii$) there exists a path $\pi'$ of $\G_A$ 
such that $\src(\pi) = \<q_0, q_0>$ for some $q_0 \in Q$, 
$\trg(\pi) = \< q_1, p, q_2>$,
and $\weight(\pi) = \bigoplus_\Deltac b_\top(q_1, p, q_2)$.
\end{lemma}
%
\begin{proof}
By associativity, commutativity and distributivity for $\Semiring$,
\eqref{eq:bbot} can be rewritten into the form, unfolding~\eqref{eq:SWVPA-weight}:
%
\begin{equation}\label{eq:bbot-unfold}
b_\bot(q, q') = 
\bigoplus_{t\in \Delta^*} 
\bigoplus_{\begin{array}{c}
           \scriptstyle q_0,\ldots, q_n \in Q\\
           \scriptstyle p_{i},\ldots, p_{k} \in P
           \end{array}}
\bigotimes_{i=1}^n w_i(\tau_i)           
\end{equation}
%
where $n$ is the length of $t$, $k \leq n$,
$q_0 = q$, $q_n = q'$, 
for all $1 \leq i < n$, 
$w_i$ is one of the functions of $\bar\wei$,
$\tau_i$ is a transition of $A$ and
$\src(\tau_i) = q_{i-1}$, $\snd(\tau_i) = q_i$.
%
Since $\Semiring$ is total, there exists finite sequences as above such that
\( b_\bot(q, q') = \bigotimes_{i=i}^n w_i(\tau_i)\).
%
There might exists, for $q$ and $q'$,
several finite sequences~$\bar\tau$ and~$\bar{w}$ as above, 
let us choose arbitrarily one of minimal length~$n$. 
This integer $n$ will be denoted $n_{q, q'}$ in the following.
%We represent a computation of $A$ on $t$ by a sequence $\rho$ of its transitions,

Similarly, following~\eqref{eq:btop} and~\eqref{eq:SWVPA-weight}, 
$\bigoplus_\Deltac b_\top(q, p, q')$ can be put in the form:
%
\begin{equation}\label{eq:btop-unfold}
{\textstyle\bigoplus_\Deltac} b_\top(q, p, q') = 
{\textstyle\bigoplus_\Deltac} \bigoplus_{t\in \Delta^*} 
\bigoplus_{\begin{array}{c}
           \scriptstyle q_0,\ldots, q_n \in Q\\
           \scriptstyle p_{i},\ldots, p_{k} \in P
           \end{array}}
\bigotimes_{i=1}^n w_i(\tau_i)           
\end{equation}
%
and hence 
\( {\textstyle\bigoplus_\Deltac} b_\top(q, p, q') = \bigotimes_{i=i}^n w_i(\tau_i) \),
using $c \in \Deltac$ that minimize the function in rhs of~\eqref{eq:btop-unfold}.
We denote the smallest $n$ as above by $n_{q, p, q'}$.

\noindent
We show now the existence of a path $\pi$ of Lemma~\ref{lem:algo-completeness}, 
by simultaneous induction on $n_{q, q'}$ and $n_{q, p, q'}$.

The base case, $n = 0$ corresponds to $t = \varepsilon$. 
In this case, by~\eqref{eq:SWVPA-weight}, $b_\top(q, p, q') = \zero$,
$b_\bot(q, q) = \one$ and $b_\bot(q, q') = \zero$ if $q \neq q'$.

\noindent
For $n = n_{q, q'} > 0$, let $\bar\tau$ and~$\bar{w}$
be the sequences associated to $q$ and $q'$ as above.
Let us do a case analysis of $w_n$.

\end{proof}
